Should I Airbnb My Property?

### 2017-01-09 abbReturns_Writeup.Rmd
Lit rev, rewrite beg, turn off analysis

---
title: "Should I AirBNB My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "December 17, 2016"
output:
  pdf_document: default
  html_document: default
---

@@

```{r load_workspace, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

@@

## Intro

With the rise of Airbnb and other self-hosting accommodation platforms, property owners and investors may consider short-term rental as a potential use for a home.  The question to be asked then is, in which circumstances would leasing in a short-term manner produce greater returns than a traditional long-term rental?  This work presents an examination of market transactions of both short and long-term rental aimed at answering this question.

To do so, we use long-term rental data and nightly Airbnb listing information from the Melbourne (Australia) metropolitan area.  Melbourne has a thriving Airbnb market with over 17,000 total properties listed on the portal.  The Airbnb data does not include exact addresses and therefore no direct matches between the datasets can be made.  As a result, we impute likely short and long-term rental rates and occupancies.  Comparisons of actual to estimated returns are made to determine the situation in which a short term rental, such as Airbnb, is more profitable.

Additionally, we break the results down by property types (apartment vs house), product type (ex. 1 bed, 1 bath) and by location (submarket).  Our results shows that the feasibility of a greater return through a short term rental is dependent on property type, product and location.  In answer to our question: it depends.

## Lit Review

@@

### Scenario

We approach answering the question of Airbnb-ing or not by examining a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne on Sept 1, 2015.
* They are looking to maximize the revenue over the 12 month period from Sept 1 2015 to August 31 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, Airbnb type rental.
* Only direct monetary costs are considered, others such as liquidity and inconvience are not considered.

We will not, a priori, specify a location or property type as we hypothesize that the feasibility of a short term rental will vary by these characteristics.  Australia has a tax policy, Negative Gearing, that does allow for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties.  However, as this policy requires knowledge of the investor's personal income level, we will not be considering this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends.  As the future trends are unknown at the time of purchase and the decision of long vs short term rental, we make our comparisons using a fixed nightly short term rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long and short term rates and occupancy structures will be baesd on the market conditions as of this date.

### Equation

In this above scenario we compare the 12-month returns from the a long term rental with those of a short term rental.  We define returns here as the gross revenue minus the direct costs associated with each of the two tenure possibilities.

$$ R_{LT} = Rev_{LT} - C_{LT} $$
$$ R_{ST} = Rev_{ST} - C_{ST} $$

$$ Rev_{LT} = Rent_{M} * (12 - C_{S}) $$  
where $C_{S}$ are search costs defined in terms of the time (in months) spent to find a long-term tenant.

$$ Rev_{ST} = Rate_{N} * O_{Total} $$

where $O_{Total}$ are the total number of occupancies over the 12-month period.

$$ C_{LT} = T + M $$

where

where $T$ are tax related expenses and $M$ are general maintenance expenses.

$$ C_{ST} = T + M + J $$

where $T$ are tax related expenses, $M$ are general maintenance expenses and $J$ are janitorial expenses incurred after each short term tenancy.

We assume for the sake of this scenario that the tax $T$ and general maintenance $M$ expenses are identical between tenure options and therefore cancel out.  With the analysis that follows we will consider short term rental to be prefable to long-term if the following inequality holds:

$$ Rate_{N} * O_{Total} - J > Rent_{M} * (12 - C_{S}) $$

@@

Data for this studies were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  The Airbnb online portal is used as evidence of the short-term rental market.  Airbnb is the largest short-term rental hub in the Melbourne region, though others do exist.  As opposed to similar sites such as VRBO.com and homeaway.com which focus on vacation homes in tourist dominated areas such as beaches and ski resorts, Airbnb offers considerable coverage in urban areas and, in way ways, directly compete with more traditional long term rentals.  The data on the AirBNB market in Melbourne have been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The short term data include two types of observations, property level and nightly.  Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year.  This data indicate the advertised price, the occupancy status and a unique reservation identification (useful for determining the length of stay).

The Airbnb data cover the Melbourne metropolitan region from October 1st 2014 to August 30th 2016. There are missing nightly observations during the time period before September 1 2015.  While we initially imputed these observations we found the results unreliable and have structured the analysis to only cover the time period containing complete data.  All monetary values in the Airbnb data are in US dollars and have been converted to Australian dollars using the median of the exchange rate during this time period, 1.32 Australian dollars to 1 US dollar.

The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company. This data also includes an observation for each rental property each time a detail of the listing changed.  For example, if a property were listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. In each case the last observation signifies the final price of the rental contract.  We separate this data into two sets for efficiency reasons. The first data set includes all information about the property, physical and economic (date and amount of rent contract), while the second contains a property ID, a rental event ID, the list price and the date of the change.  As such, the first dataset is a set of market observations with associated characteristics and the second is a listing history of each event. The long-term lease data covers the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two dataset as the Airbnb data does not include address.  Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates.  As a result of this, we are unable to directly match observations from the two datasets.  However, the number of properties that rented long term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  exch.rate <- 1.32

### Preparation

In this section we describe the operations that we have performed to prepare the two sets of raw data for analysis.  Primarily, this involves the filtering out of data errors and/or outlier values.  However, as our data are from two different sources, number of fields also require standardization so that comparisons between the two tenure options can be made.

We begin by filtering observations by time. Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term rental versus an AirBNB, short term rental approach to generating income. To determine the likely long term rental rate for any given property, we will use the observed rental transactions from the market from September 1, 2014 to August 31, 2015. Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.

Next, we examine property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

1. **House**:  Includes properties labeled 'House' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Apartment' or 'Condominium'
3. **Other**: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Loft', 'Other', 'Tent', 'Treehouse', Villa' and 'Yurt.

Within the long-term rental data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

1. **House**:  Includes properties labeled 'Duplex', 'House', 'Terrace' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Unit' or 'Studio'
3. **Other**: Includes properties labeled as 'Semi' and 'Villa'

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

As the chart below show, the **Other** category only makes up a very small percentage of the properties in both the short and long term market, especially in the long-term market.  Additionally, a good portion of the **Others** in the Airbnb data are Bed & Breakfast units, a use which doesn't fit with our research question.  As a result, we remove the 'Others' from both the Airbnb and the long term rental data.

```{r load_beforePT, message=FALSE, warning=FALSE, cache=TRUE, comment=FALSE, echo=FALSE, results='hide'}  
  data.path <- 'c:/dropbox/research/airBNB/data/'
  load(paste0(data.path, '/analyzed/abb_beforePT.RData'))
  abb.data <- before.PT$abb.data
  ltr.data <- before.PT$ltr.data
  daily.data <- before.PT$daily.data

```{r type_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.cap='Tenure Type by Property Type'}

  # Create Airbnb plot

  abb.types <- ggplot(abb.data, aes(x=type)) + 
    geom_bar(color=str.col, fill=str.col) +
    xlab('') + 
    ylab('# of Properties\n') +
    labs(title = "Short Term Rentals\n") + 
    theme(plot.title = element_text(hjust = 0.5))

  # Create long term rental plot

  ltr.types <- ggplot(ltr.data, aes(x=type)) + 
    geom_bar(color=ltr.col, fill=ltr.col) +
    xlab('') + 
    ylab('# of Properties\n') +
    labs(title = "Long Term Rentals\n") + 
    theme(plot.title = element_text(hjust = 0.5))

  # Output plots

  ggMultiPlots(abb.types, ltr.types, cols=2)

```{r remove_others, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

  abb.data <- subset(abb.data, type != 'Other')
  ltr.data <- subset(ltr.data, type != 'Other')

Next, for the short term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked:

1. **Entire Home/Apt**: The entire home or apartment is available
2. **Private Room**: One room within a house or apartment is available
3. **Shared Room**: A bed within a room shared by another occupant(s) is available

As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accomodations and other types of rental which may offer private or shared room options] and our purpose here is a comparison of long and short term returns, we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.

```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap='Listing Type of Short Term Rentals'}

  # Fix data types and names

  abb.data$listing.type <- as.character(abb.data$listing.type)
  abb.data$listing.type[abb.data$listing.type == 'Entire home/apt'] <- 'Entire home'

  # Make Plot

  ggplot(abb.data, aes(x=listing.type)) + 
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('') + 
    ylab('# of Properties\n') +
    labs(title = "Short Term Rentals by Listing Type\n") + 
    theme(plot.title = element_text(hjust = 0.5))

```{r keep_entire_home, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

  abb.room <- abb.data[abb.data$listing.type != 'Entire home', ]
  abb.data <- abb.data[abb.data$listing.type == 'Entire home', ]

#### Structural Characteristics

We now turn to filtering by structural characteristics.  As the Airbnb data provides limited information on the structural characteristics, the only overlapping fields that we have between the two datasets are bedrooms and bathrooms. Additionally, as the markets and structural components of houses and apartment differ, we bifurcate our analysis between the two property types from this point forward.

We begin by removing all properties that are missing bedroom information or have a likely data error in this fields (more than 14 bedrooms). For both short and long term rentals we see that most apartments have less than 4 bedrooms and houses less than 5.  We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units.

```{r bed_fltr, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

 # Remove missing and data errors

  # Long term data
  ltr.data <- ltr.data[!is.na(ltr.data$bedrooms), ]
  ltr.data <- ltr.data[ltr.data$bedrooms <= 10, ]

  # Short term data
  abb.data <- abb.data[!is.na(abb.data$bedrooms), ]
  abb.data <- abb.data[abb.data$bedrooms <= 10, ]

```{r bed_fltr2, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.height=3.5, fig.cap='Bedrooms by Tenure & Property Type'}

 # Build long term bar plot

   ltrbed <- ggplot(ltr.data, aes(x=bedrooms)) + 
    geom_bar(fill=ltr.col, color=ltr.col) +
    facet_wrap(~type) +
    scale_x_continuous(labels=1:10, breaks=1:10)+
    xlab('') +
    ylab('# of Props\n') +
    labs(title = "Bedrooms in Long Term Rentals\n") + 
    theme(plot.title = element_text(hjust = 0.5))

 # Build short term bar plot

  abbbed <- ggplot(abb.data, aes(x=bedrooms)) + 
    geom_bar(fill=str.col, color=str.col) +
    facet_wrap(~type) +
    scale_x_continuous(labels=0:10, breaks=0:10)+
    xlab('') +
    ylab('# of Props\n') +
    labs(title = "Bedrooms in Short Term Rentals\n") + 
    theme(plot.title = element_text(hjust = 0.5))

 # Make the comparison plot

  ggMultiPlots(abbbed, ltrbed, cols=1)

```{r bed_fltr3, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

 # Remove all with more than 4 beds

>>>>

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "January 9, 2017"
output:
  pdf_document: default
  html_document: default
geometry: margin=3.4cm
bibliography: abb.bib
---

@@

```{r load_workspace, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, cache=TRUE}

@@

&nbsp;
&nbsp;

## Introduction

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential use for a revenue generating residential property. Should they though?  And which types of properties and in which locations would short-term leasing produce greater revenue than a traditional long-term rental (LTR), if any?

To address these questions and provide some guidance to potential and current investors and owners, we use long-term rental data and nightly Airbnb listing information from the Melbourne (Australia) metropolitan area.  Melbourne has a thriving Airbnb market with over 20,000 total properties listed on the portal. We adopt a scenario-based approach in which we compare potential annual revenues for properties under both a short-term rental (STR) and a long-term rental (LTR) situation. Additionally, we break the results down by property types (apartment vs house), product type (ex. 1 bed, 1 bath) and by location (submarket).  Our results shows that the feasibility of a greater return through a short term rental is dependent on property type, product and location.  In answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

## Literature Review

Existing research into short-term rental markets is sparse. However, with the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields outside of real estate as evidenced by a growth in recent publications and on-going working papers.

Much of the existing work focuses on price or rent formation in second home markets.  Specifically, the impacts of amenties such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water] and proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  @benjamin2001value note that leasing policies, in this case smoking prohibitions, can also influence the weekly rental rates for vacation homes.  @nilsson2015influence makes a key distinction, finding that rural vacation home markets value amenties whereas urban holiday properties tend to derive large premiums from proximity to attractive residential areas.

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast, the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability toward owners.  The authors also note that very little research has been put toward understanding this growing market segment.

A set of recent work has looked into the impact of short-term markets on housing prices and affordability. These working papers examine the influence of Airbnb on existing housing prices in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2016home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other work examines the impact of short term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

A critical theme that arises out of the extant literature is that fundamentals of the markets for primary dwellings (owner-occupied and long-term rentals) may differ from those of vacation homes (or, by association, short term rentals). Or, in other words, the factors that influence prices and rents in the two are not equivalent. @cho2003impacts find that the second home market is more sensitive to proximity to environmental amenities than the owner occupied market. In Switzerland, @soguel2008impact test to see if second homes pay a higher premium for ski resort access.  Their results fail to show a statistically significant difference in the impact of ski resorts between markets, but do find that other property characteristics have heterogenous effects between the two.

Though issues of second homes were primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has re-focused some of the problems of short term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short term rentals, particulary though Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, impacts on housing market and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy. The closest analog in the literature is a piece by @larsen2009purchase examining the choice between purchasing a time-share or simply paying for hotel rooms.  Theoretically, the literature shows that in most cases there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that this same thinking can be extended to the difference between short-term occupants (tourists) and long-term rental occupants (workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term tenure.

@@

We approach answering the question of whether to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne on Sept 1, 2015.
* They are looking to maximize the revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, Airbnb type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvience are not considered.

We will not, a priori, specify a location or property type as we hypothesize that the preference of a short-term rental over a long-term one will vary by the location and property characteristics.  Additionally, Australia implements a particular tax policy, often termed `negative gearing', that does allow for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties^[See @fane2004negative for a complete explanation of negative gearing.  However, as the negative gearing policy requires knowledge of the investor's personal income situation, we will not be considering any effects of this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long- term rents and short-term rates and occupancy structures will be baesd on the market conditions as of this date.

We do not directly observe properties that are utilized in both a short and long term leasing strategy.  Therefore, comparisons are done based on estimated or imputed revenues.  Most specifically, we will estimate short-term revenue for observations of long-term rentals (LTRs) and long-term revenue for observations of short-term rentals (STRs). A preference for short or long term leasing strategy will be determined based on comparing the observed and estimated revenues.

### Comparison

In this above scenario we compare the 12-month revenues from the a long-term rental with those of a short-term rental.  We define returns here as the gross revenue minus the direct costs associated with each of the two tenure possibilities.

\begin{equation}
Rev_{LT} = RevG_{LT} - C_{LT}
\end{equation}

\begin{equation}
Rev_{ST} = RevG_{ST} - C_{ST}
\end{equation}

Where $Rev_{LT}$ is the net revenue for a long-term rental, $RevG_{LT}$ is the gross revenue and $C_{LT}$ are the costs associated with long-term rental. The gross long-term revenue is expressed by:

\begin{equation}
RevG_{LT} = Rent_{M} * (12 - C_{S})
\end{equation}

where $Rent_{M}$ is the nominal monthly rent and $C_{S}$ are search costs defined in terms of the time (in months) spent to find a long-term tenant.

The short-term revenue is expresse by:

\begin{equation}
RevG_{ST} = Rate_{N} * O_{Total}
\end{equation}

where $Rate_{N}$ is the average nightly rate and $O_{Total}$ are the total number of occupancies (bookings) over the 12-month period.

We have limited our consideration of costs in this case to be those undergone during a single year of ownership.  Broadly speaking, these fall into three categories for long-term rental and four categories for short-term properties.  Long-term costs are summarized as:

\begin{equation}
C_{LT} = T + M + F
\end{equation}

where $T$ are tax related expenses, $M$ are general maintenance expenses and $F$ are finaning costs. Short-term properies experience the same three costs ($T$, $M$ and $F$) plus additional janitorial costs between each of the short-term occupants.

\begin{equation}
C_{ST} = T + M + F + J
\end{equation}

As our scenario compares a hypothetical and particular owner over the course of a single  year the tax $T$ and $F$ costs are identical between the two leasing strategy and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliance and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.].  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short or long term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

Finally, we have janitorial costs $J$. These cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experiences every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb. Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior.  As a result of janitorial fees being a separate line item and directly paid by the occupants can also be ignored in the comparative analysis below as they are direct.

Given that the costs cancel out between the two options, we will consider short-term rental (STR) strategy to be prefable to long-term rental (LTR) strategy if the following inequality holds:

\begin{equation}
RevG_{ST} > RevG_{LT} 
\end{equation}

More specifically, if:

\begin{equation}
Rate_{N} * O_{Total} > Rent_{M} * (12 - C_{S})
\end{equation}

@@

<!-- Data for this studies were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  The Airbnb online portal is used as evidence of the short-term rental market.  Airbnb is the largest short-term rental hub in the Melbourne region, though others do exist.  As opposed to similar sites such as VRBO.com and homeaway.com which focus on vacation homes in tourist dominated areas such as beaches and ski resorts, Airbnb offers considerable coverage in urban areas and, in way ways, directly compete with more traditional long term rentals.  The data on the AirBNB market in Melbourne have been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The short term data include two types of observations, property level and nightly.  Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year.  This data indicate the advertised price, the occupancy status and a unique reservation identification (useful for determining the length of stay). -->

<!-- The Airbnb data cover the Melbourne metropolitan region from October 1st 2014 to August 30th 2016. There are missing nightly observations during the time period before September 1 2015.  While we initially imputed these observations we found the results unreliable and have structured the analysis to only cover the time period containing complete data.  All monetary values in the Airbnb data are in US dollars and have been converted to Australian dollars using the median of the exchange rate during this time period, 1.32 Australian dollars to 1 US dollar.  -->

<!-- The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company. This data also includes an observation for each rental property each time a detail of the listing changed.  For example, if a property were listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. In each case the last observation signifies the final price of the rental contract.  We separate this data into two sets for efficiency reasons. The first data set includes all information about the property, physical and economic (date and amount of rent contract), while the second contains a property ID, a rental event ID, the list price and the date of the change.  As such, the first dataset is a set of market observations with associated characteristics and the second is a listing history of each event. The long-term lease data covers the period from January 2014 to December 2015.   -->

<!-- Unfortunately, there are no common identifiers between the two dataset as the Airbnb data does not include address.  Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates.  As a result of this, we are unable to directly match observations from the two datasets.  However, the number of properties that rented long term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.  -->

<!-- ```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'} -->

<!--   exch.rate <- 1.32   -->

<!-- ``` -->

<!-- ### Preparation -->

<!-- In this section we describe the operations that we have performed to prepare the two sets of raw data for analysis.  Primarily, this involves the filtering out of data errors and/or outlier values.  However, as our data are from two different sources, number of fields also require standardization so that comparisons between the two tenure options can be made.  -->

<!-- We begin by filtering observations by time. Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term rental versus an AirBNB, short term rental approach to generating income. To determine the likely long term rental rate for any given property, we will use the observed rental transactions from the market from September 1, 2014 to August 31, 2015. Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.   -->

<!-- Next, we examine property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types: -->

<!-- 1. **House**:  Includes properties labeled 'House' or 'Townhouse' -->
<!-- 2. **Apartment**: Includes properties labeled as 'Apartment' or 'Condominium' -->
<!-- 3. **Other**: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Loft', 'Other', 'Tent', 'Treehouse', Villa' and 'Yurt.  -->

<!-- Within the long-term rental data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories: -->

<!-- 1. **House**:  Includes properties labeled 'Duplex', 'House', 'Terrace' or 'Townhouse' -->
<!-- 2. **Apartment**: Includes properties labeled as 'Unit' or 'Studio' -->
<!-- 3. **Other**: Includes properties labeled as 'Semi' and 'Villa' -->

<!-- Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.  -->

Berry 2000 about unique australia market

<!-- As the chart below show, the **Other** category only makes up a very small percentage of the properties in both the short and long term market, especially in the long-term market.  Additionally, a good portion of the **Others** in the Airbnb data are Bed & Breakfast units, a use which doesn't fit with our research question.  As a result, we remove the 'Others' from both the Airbnb and the long term rental data.  -->

<!-- ```{r load_beforePT, message=FALSE, warning=FALSE, cache=TRUE, comment=FALSE, echo=FALSE, results='hide'}   -->
<!--   data.path <- 'c:/dropbox/research/airBNB/data/' -->
<!--   load(paste0(data.path, '/analyzed/abb_beforePT.RData')) -->
<!--   abb.data <- before.PT$abb.data -->
<!--   ltr.data <- before.PT$ltr.data -->
<!--   daily.data <- before.PT$daily.data -->

<!-- ```  -->

<!-- ```{r type_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.cap='Tenure Type by Property Type'} -->

<!--   # Create Airbnb plot -->

<!--   abb.types <- ggplot(abb.data, aes(x=type)) +  -->
<!--     geom_bar(color=str.col, fill=str.col) + -->
<!--     xlab('') +  -->
<!--     ylab('# of Properties\n') + -->
<!--     labs(title = "Short Term Rentals\n") +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--   # Create long term rental plot -->

<!--   ltr.types <- ggplot(ltr.data, aes(x=type)) +  -->
<!--     geom_bar(color=ltr.col, fill=ltr.col) + -->
<!--     xlab('') +  -->
<!--     ylab('# of Properties\n') + -->
<!--     labs(title = "Long Term Rentals\n") +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--   # Output plots -->

<!--   ggMultiPlots(abb.types, ltr.types, cols=2) -->

<!-- ``` -->

<!-- ```{r remove_others, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE} -->

<!--   abb.data <- subset(abb.data, type != 'Other') -->
<!--   ltr.data <- subset(ltr.data, type != 'Other') -->

<!-- ``` -->

<!-- Next, for the short term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked: -->

<!-- 1. **Entire Home/Apt**: The entire home or apartment is available -->
<!-- 2. **Private Room**: One room within a house or apartment is available -->
<!-- 3. **Shared Room**: A bed within a room shared by another occupant(s) is available -->

<!-- As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accomodations and other types of rental which may offer private or shared room options] and our purpose here is a comparison of long and short term returns, we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.  -->

<!-- ```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap='Listing Type of Short Term Rentals'} -->

<!--   # Fix data types and names -->

<!--   abb.data$listing.type <- as.character(abb.data$listing.type) -->
<!--   abb.data$listing.type[abb.data$listing.type == 'Entire home/apt'] <- 'Entire home' -->

<!--   # Make Plot -->

<!--   ggplot(abb.data, aes(x=listing.type)) +  -->
<!--     geom_bar(color=abb.col[1], fill=abb.col[1]) + -->
<!--     xlab('') +  -->
<!--     ylab('# of Properties\n') + -->
<!--     labs(title = "Short Term Rentals by Listing Type\n") +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!-- ``` -->

<!-- ```{r keep_entire_home, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE} -->

<!--   abb.room <- abb.data[abb.data$listing.type != 'Entire home', ] -->
<!--   abb.data <- abb.data[abb.data$listing.type == 'Entire home', ] -->

<!-- ``` -->

<!-- #### Structural Characteristics -->

<!-- We now turn to filtering by structural characteristics.  As the Airbnb data provides limited information on the structural characteristics, the only overlapping fields that we have between the two datasets are bedrooms and bathrooms. Additionally, as the markets and structural components of houses and apartment differ, we bifurcate our analysis between the two property types from this point forward.  -->

<!-- We begin by removing all properties that are missing bedroom information or have a likely data error in this fields (more than 14 bedrooms). For both short and long term rentals we see that most apartments have less than 4 bedrooms and houses less than 5.  We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units. -->

<!-- ```{r bed_fltr, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE} -->

<!--  # Remove missing and data errors -->

<!--   # Long term data -->
<!--   ltr.data <- ltr.data[!is.na(ltr.data$bedrooms), ] -->
<!--   ltr.data <- ltr.data[ltr.data$bedrooms <= 10, ] -->

<!--   # Short term data -->
<!--   abb.data <- abb.data[!is.na(abb.data$bedrooms), ] -->
<!--   abb.data <- abb.data[abb.data$bedrooms <= 10, ] -->

<!-- ``` -->
